Coherent Hyper-Raman Four Wave Mixing Spectroscopies

### 2024-06-11 paper.tex
fixed placzek citation
edits
more on quantitative stuff.

\documentclass[aip, jcp, reprint, onecolumn]{revtex4-2}

@@

It is useful to expose relationships between transition dipoles, Raman polarizabilities and hyper-Raman polarizabilities in the driven limit. \cite{Simpson2004, RN120}
In the driven limit, under the electric dipole approximation, the I$^{th}$ component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of any four wave mixing process, induced by electric fields E, at output frequency $\omega_4$ is written as \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J E_K E_L 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the IJKL element of the third order electrical susceptibility, a rank four tensor. 
Note that we choose the convention where an arbitrary electric field, assuming a plane wave, is written in Fourier components as 
\begin{equation}
	\vec{E}(\omega) = \frac{1}{2} E_0 \exp(i(\vec{k}\cdot\vec{x} - \omega t)) + c.c.
\end{equation}
where $E_0$ is the pulse amplitude $\vec{k}$ is a wavevector.
The $IJKL$ element of $\chi^{(3)}$, $\chi^{(3)}$ is generally described as
\begin{equation}
	\chi^{(3)}_{IJKL} = NF(\omega_4) \langle \gamma_{ijkl} \rangle
\end{equation}
where N is a number density, F is the Lorentz local field factor, and $\gamma_{ijkl}$ is the third order polarizability (i.e., second hyperpolarizability). 
The brackets indicate an orientational average. 
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to molecular frame coordinates.

Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re{\chi^{(3)}_{IJKL}}$ and $\Im{\chi^{(3)}_{IJKL}}$.\cite{Levenson1974_1, Levenson1974_2}
Resonant SIVE lineshapes are complicated by amplitude level interference between the sample and its substrate and/or sample cell windows. \cite{RN362}
These effects have been treated in excellent detail for three laser infrared-infrared-visible SIVE using the pathway in \autoref{fig:sivewmel}a, and provide methods to export quantitative information from SIVE spectra. \cite{RN418}

@@

\subsection{SR-SIVE Selection Rules}
To make the connection between SR-SIVE and hyper-Raman scattering, we investigate the gross selection rules of SR-SIVE-IR.
By propagating density matrix elements in the steady state limit, the SIVE hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{i}_{v n} \mu^{j}_{nm} \mu^{k}_{mg} \mu^{l}_{gv} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{j}_{ab}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state density.
For simplicity, we consider the results for the process diagrammed in \autoref{fig:sivewmel}a, but by multiplying by (-1) and taking a complex conjugate, which arise from the bra-side transition in \autoref{fig:sivewmel}a, results for the \autoref{fig:sivewmel}b  process are obtained.
Note that $\omega_j \rightarrow -\omega_j$ for a bra side transition.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$ and is defined as $\frac{n!}{(n-m)!}$, where $n$ is the total number of input fields (3 for any four wave mixing experiment) and $m$ is the number of distinguishable input fields.\cite{RN134} 
For a SIVE experiment using two or three distinct input fields, as considered here, D = 6.
Note that $\ket{m}, \ket{n}$ are virtual states (\autoref{fig:sivewmel}).
By contracting over the virtual states to form the hyper-Raman hyperpolarizability $\beta$ (i.e., Placzek approximation), this expression can be simplified immensely.\cite{Long1970} 
Note, however, that if $\omega_2$, $\omega_3$, as depicted in \autoref{fig:sivewmel}, are nearly resonant with any state, the Placzek approximation fails. \cite{Placzek1937, Long1970}
Assuming the validity of the Placzek approximation, \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{(ijk)}_{gv} \mu^{(l)}_{gv}}{\Delta_{gv}} \rho_{gg}
\end{equation}
As such, any mode that is SIVE active must be both hyper-Raman and IR active.
However, since any IR active vibrational mode is hyper-Raman active,\cite{Cyvin1965, Andrews1978} any IR active mode is SIVE active.
This makes SIVE an optically upconverted infrared spectroscopy of isotropic samples, as its selection rules are identically those of IR spectroscopy, but its output is generally in the visible.
Unlike SFG spectroscopy, which is dependent upon both the infrared and Raman activity of a mode (through symmetry constraints, this demands that the microscopic SFG output from centrosymmetric species vanishes),\cite{Shen90, Cotton} SIVE is a nonlinear, optically upconverted spectroscopy for any infrared active mode.

@@

To motivate the application of SIVE spectroscopy, we perform an approximate calculation to compare the output of parametric SIVE-IR (\autoref{fig:sivewmel}b) to that of vibrational SFG (vSFG) spectroscopy.
Absorption effects are neglected for simplicity.
To allow comparison between the two methods, the output polarizations of each process is written as
\begin{widetext}
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{SIVE}}} = N_{bulk}  F(\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(\omega_1)} \l_{eff}
		\end{equation}
	\end{subequations}
\end{widetext}
where $\alpha_{ij}$ is the Raman polarizability, N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We assume the bulk optical interaction length ($\l_{eff}$) to be roughly 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored and the input and output fields are all co-polarized (I=J=K=L=Z), so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{SIVE}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \l_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{Ziegler87_CPL, RN515}
Such an intensity is easily obtained using modern ultrafast sources.
As such,
\begin{equation}
\begin{split}
		P_{ratio} &\sim 10^4 \frac{\beta E(\omega_3)}{\alpha} \sim 10
\end{split}
\end{equation}
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the SIVE output is roughly 100 times stronger than vibrational SFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

\subsection{Interference in SR-SIVE}
It is well known that nonlinear wave mixing processes which share the same time ordering and output frequency will interfere. \cite{RN135, Bonn2024}
Understanding interfering pathways is useful because they can, in some cases, eliminate output. \cite{RN342, RN135}
The WMEL diagrams presented in \autoref{fig:sivewmel} outline resonant SIVE processes when the infrared pulse $(\omega_1)$ is resonant with a vibrational mode. 
However, if $\omega_2$ becomes vibrationally resonant and $\omega_1$ is non-resonant, then other SIVE pathways appear.\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:sivewmel}, the WMEL diagrams presented in \autoref{fig:sivewmel2} may significantly interfere with DOVE-IR and DOVE-Raman spectra.
The time ordering in DOVE-IR and DOVE-Raman spectra is used to interpret anharmonicities, as DOVE-IR and DOVE-Raman destructively interfere when outputs are at the same frequency. \cite{RN135, RN324}
It is therefore important to understand how the corresponding SIVE pathways interfere.
To this end, we investigate how the SIVE-IR-II and SIVE-Raman pathways (\autoref{fig:sivewmel2}a, b, respectively) interfere under the rotating wave approximation.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{main/wmel2.png}
	\caption{Wave Mixing Energy Level (WMEL) diagrams of interfering SR-SIVE pathways in the  $\vec{k}_4 = -\vec{k}_1 + \vec{k}_2 + \vec{k}_3$ phasematching geometry. 
	(a) is similar to the process diagrammed in \autoref{fig:sivewmel}, whereas (b) is similar to a singly resonant coherent anti-Stokes Raman-type process.}
	\label{fig:sivewmel2}
\end{figure}

We employ the response function formalism of Mukamel for this analysis. \cite{RN287}
The third order response functions, $R^{(3)}$, for SR-SIVE-IR-II ($R^{(3)}_{S_{II}}$) and SR-SIVE-Raman ($R^{(3)}_{R}$) (\autoref{fig:sivewmel2}a, b) are 
\begin{subequations}
	\begin{equation} \label{mixing:a}
		R^{(3)}_{S_{II}} (t_2) = -\frac{i}{\hbar} [\beta(t_2), \mu(0)] \exp(-i\Delta_{v''g}t_2) \theta(t_2)
	\end{equation}
	\begin{equation}\label{mixing:b}
		\begin{split}
			R^{(3)}_{R} (t_2) & = \frac{i}{\hbar} [\beta(t_2), \mu(0)] \exp(-i\Delta_{v''g}t_2) \theta(t_2) = - R^{(3)}_{S_{II}} (t_2)\\
		\end{split}
	\end{equation}
\end{subequations}
By assuming that the beams which stimulate hyper-Raman scattering, $\vec{E}_1, \vec{E}_3$, interact simultaneously at some time $\tau$ after $\vec{E}_2$ interacts with the sample, the total output polarization, $\vec{P}^{(3)}(t)$, is written as \cite{Cho2001}
\begin{widetext}
	\begin{equation}\label{polarization_SRSIVE_a}
		\begin{split}
			\vec{P}^{(3)}(t) &= \vec{P}_{S_{II}}(t) + \vec{P}_{R}(t)\\
			&= \int_0^\infty \mathrm{d}t_2 (R^{(3)}_{S_{II}} (t_2) + R^{(3)}_{R} (t_2))\vdots \vec{E}_3(t) \vec{E}_1(t) \vec{E}_2(t-t_2 + \tau)\\
			&= \int_0^\infty \mathrm{d}t_2 (R^{(3)}_{S_{II}} (t_2) - R^{(3)}_{S_{II}} (t_2))\vdots \vec{E}_3(t) \vec{E}_1(t) \vec{E}_2(t-t_2 + \tau) \\
			&= 0
		\end{split}
	\end{equation}
\end{widetext}
It is clear the total output polarization for SR-SIVE-IR-II and SR-SIVE-Raman vanishes, or, SR-SIVE-IR-II and SR-SIVE-Raman perfectly destructively interfere.
The absence of SR-SIVE-IR-II pathways in ultrafast DOVE work is rationalized by this calculation. \cite{RN367, McDonnell2024}
In the case of the time ordered SR-SIVE-IR-I pathway, there are no opportunities for interference as all other pathways are nonresonant with the first interaction. 
A possible mechanism for interfering output in the SR-SIVE-I pathway are cascaded second order processes (SFG, DFG). \cite{RN243, RN300}
In isotropic media, SFG and DFG vanish under the electric dipole approximation,\cite{RN231} so cascades are unlikely to inhibit output in centrosymmetric media.
In media where SFG and DFG are allowed, e.g. surfaces or noncentrosymmetric media, second order cascades may become important. 
Generally, for nonlinear spectroscopies which share the same phasematching conditions as SR-SIVE (e.g., DOVE), it is useful to note that SR-SIVE-IR-I is not canceled, whereas SR-SIVE-IR-II and SR-SIVE-Raman permanently cancel.

\subsection{Rovibrational SR-SIVE Selection Rules} %this can probably be deleted but wanted to add just in case...
Recent developments in nonlinear spectroscopy have demonstrated the feasibility of multidimensional infrared spectroscopy of gaseous and quasi-free species. \cite{Ziegler2018, Gronborg2022, RN325, Chen2024}
DOVE and 2D-IR have been used to isolate coupling between fundamentals, overtones and combination bands of gaseous species in several molecular systems. 
Since resonance hyper-Raman scattering spectroscopy was developed as a gas phase spectroscopy, we briefly investigate the properties of rovibrational SR-SIVE.\cite{Ziegler1987}

The rovibrational selection rules are generated through calculating \autoref{sivebeta} in a $\ket{vJM} \equiv \ket{v} \otimes {\ket{JM}}$ basis, where $\{\ket{v}\}$ is a basis of vibrational states and $\{\ket{JM}\}$ is a basis of rotational states ($J$ is the rotational quantum number and $M$ is its z projection, respectively).
By writing $\vec{\mu} = T_{q_1}^{(1)}$ and $\beta_{ijk} = T_{q_1}^{(1)} + T_{q_2}^{(2)} + T_{q_3}^{(3)}$, where $T^{(k)}_{q}$ is a rank k spherical tensor, it is simple to evaluate the integrals in \autoref{sivebeta} with the Wigner-Eckart theorem. \cite{Kowzan2022}
It is important to note that $\beta_{ijk}$ does not have complete tensor index symmetry.
Specifically, it has shown that the rank one spherical tensor ($T_{q_1}^{(1)}$, with $q_1 \in -1, 0 ,1$) fully contributes to $\beta_{ijk}$ but only select $T_{q_2}^{(2)} + T_{q_3}^{(3)}$ terms contribute. \cite{Andrews1978, Andrews1990}
%there is some fancy way to note that theres no index constraint but i dont remember the language...
As such, its decomposition into irreducible tensor form is allowed only if the values of $q', q''$ are constrained.
By substituting the spherical tensor definitions of $\vec{\mu}$ and $\beta_{ijk}$ into \autoref{sivebeta}, it is clear that only $\Delta J = \pm 1$; $\Delta M = 0, \pm1$ transitions are allowed.
This is expected since $\vec{\mu}$ transitions demand $\Delta J = \pm 1$.
As a result, consistent with vibrational SR-SIVE, rovibrational SR-SIVE follows the same rules as common infrared rovibrational spectroscopy.

\subsection{DR-SIVE-IR Selection Rules}
DR-SIVE (where a vibrational and vibronic state are resonantly coupled) can provide a tool for measuring vibronic coupling.
This method is analogous to doubly resonant sum frequency generation. \cite{Shen94}
Multi-resonant nonlinear spectroscopies, such as coherent anti-Stokes Raman spectroscopy (CARS), 2D Electronic Vibrational (2D-EV) spectroscopy and triply resonant sum frequency generation (TRSF) spectroscopy, have been employed to resolve vibronic coupling in molecular samples. \cite{Carlson1990, Gaynor2017, RN276}
DR-SIVE provides methods to investigate vibronic coupling with only two laser pulses, as long as the electronic state is accessible via two photon absorption from the ground state (\textit{vide infra}).
Two beam, parametric and non-parametric DR-SIVE processes are diagrammed in \autoref{fig:doubsive}.
Here, states are written assuming the Born-Oppenheimer approximation, such that $\ket{a,b} = |a) \otimes \ket{b}$ for electronic states $\{|a)\}$ and vibrational states $\{\ket{b}\}$. \cite{BornOppenheimer, Albrecht1960}

>>>>

\documentclass[aip, jcp, reprint, twocolumn]{revtex4-2}

@@

It is useful to expose relationships between transition dipoles, Raman polarizabilities and hyper-Raman polarizabilities in the driven limit. \cite{Simpson2004, RN120}
In the driven limit, under the electric dipole approximation, the I$^{th}$ component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of any four wave mixing process, induced by electric fields E, at output frequency $\omega_4$ is written as \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J E_K E_L 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the IJKL element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}
	\chi^{(3)}_{IJKL} = NF(\omega_4) \langle \gamma_{ijkl} \rangle
\end{equation}
where N is a number density, F is the Lorentz local field factor, and $\gamma_{ijkl}$ is the third order polarizability (i.e., second hyperpolarizability). 
The brackets indicate an orientational average. 
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to molecular frame coordinates.

Note that we choose the convention where an arbitrary electric field, assuming a plane wave, is written in Fourier components as 
\begin{equation}
	\vec{E}(\omega) = \frac{1}{2} \vec{E}_0 \exp(i(\vec{k}\cdot\vec{x} - \omega t)) + c.c.
\end{equation}
where $E_0$ is the pulse amplitude and $\vec{k}$ is a wavevector.

@@

\subsection{Vibrational SR-SIVE Selection Rules}
To make the connection between SR-SIVE and hyper-Raman scattering, we investigate the gross selection rules of SR-SIVE-IR.
By propagating density matrix elements in the steady state limit, the SIVE hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{i}_{v n} \mu^{j}_{nm} \mu^{k}_{mg} \mu^{l}_{gv} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{j}_{ab}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state density.
For simplicity, we consider the results for the process diagrammed in \autoref{fig:sivewmel}a, but by multiplying by (-1) and taking a complex conjugate, which arise from the bra-side transition in \autoref{fig:sivewmel}a, results for the \autoref{fig:sivewmel}b  process are obtained.
Note that $\omega_j \rightarrow -\omega_j$ for a bra side transition.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$ and is defined as $\frac{n!}{(n-m)!}$, where $n$ is the total number of input fields (3 for any four wave mixing experiment) and $m$ is the number of distinguishable input fields.\cite{RN134} 
For a SIVE experiment using two or three distinct input fields, as considered here, D = 6.
Note that $\ket{m}, \ket{n}$ are virtual states (\autoref{fig:sivewmel}).
By contracting over the virtual states to form the hyper-Raman hyperpolarizability $\beta$ (i.e., Placzek approximation), this expression can be simplified immensely.\cite{Long1970} 
Note, however, that if $\omega_2$, $\omega_3$, as depicted in \autoref{fig:sivewmel}, are nearly resonant with any state, the Placzek approximation fails. \cite{Placzek1934, Long1970}
Assuming the validity of the Placzek approximation, \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{(ijk)}_{gv} \mu^{(l)}_{gv}}{\Delta_{gv}} \rho_{gg}
\end{equation}
As such, any mode that is SIVE active must be both hyper-Raman and IR active.
However, since any IR active vibrational mode is hyper-Raman active,\cite{Cyvin1965, Andrews1978} any IR active mode is SIVE active.
This makes SIVE an optically upconverted infrared spectroscopy of isotropic samples, as its selection rules are identically those of IR spectroscopy, but its output is generally in the visible.
Unlike SFG spectroscopy, which is dependent upon both the infrared and Raman activity of a mode (through symmetry constraints, this demands that the microscopic SFG output from centrosymmetric species vanishes),\cite{Shen90, Cotton} SIVE is a nonlinear, optically upconverted spectroscopy for any infrared active mode.

@@

To motivate the application of SIVE spectroscopy, we perform an approximate calculation to compare the output of parametric SIVE-IR (\autoref{fig:sivewmel}b) to that of vibrational SFG (vSFG) spectroscopy.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{SIVE}}} = N_{bulk}  F(\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where $\alpha_{ij}$ is the Raman polarizability, N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
\end{widetext}
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{SIVE}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
As such,
\begin{equation}
\begin{split}
		P_{ratio} &\sim 10^4 \frac{\beta E(\omega_3)}{\alpha} \sim 10
\end{split}
\end{equation}
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the SIVE output is roughly 100 times stronger than vibrational SFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

\subsection{Rovibrational SR-SIVE Selection Rules} %this can probably be deleted but wanted to add just in case...
Recent developments in nonlinear spectroscopy have demonstrated the feasibility of multidimensional infrared spectroscopy of gaseous and quasi-free species. \cite{Ziegler2018, Gronborg2022, RN325, Chen2024}
DOVE and 2D-IR have been used to isolate coupling between fundamentals, overtones and combination bands of gaseous species in several molecular systems. 
Since resonance hyper-Raman scattering spectroscopy was developed as a gas phase spectroscopy,\cite{RN515} we briefly investigate the properties of rovibrational SR-SIVE.

The rovibrational selection rules are generated through calculating \autoref{sivebeta} in a $\ket{vJM} \equiv \ket{v} \otimes {\ket{JM}}$ basis, where $\{\ket{v}\}$ is a basis of vibrational states and $\{\ket{JM}\}$ is a basis of rotational states ($J$ is the rotational quantum number and $M$ is its z projection, respectively).
By writing $\vec{\mu} = T_{q_1}^{(1)}$ and $\beta_{ijk} = T_{q_1}^{(1)} + T_{q_2}^{(2)} + T_{q_3}^{(3)}$, where $T^{(k)}_{q}$ is a rank $k$ spherical tensor, it is simple to evaluate the integrals in \autoref{sivebeta} through the Wigner-Eckart theorem. \cite{Kowzan2022}
It is important to note that $\beta_{ijk}$ does not have complete tensor index symmetry.
Specifically, it has shown that the rank one spherical tensor ($T_{q_1}^{(1)}$, with $q_1 \in \{-1, 0 ,1\}$) fully contributes to $\beta_{ijk}$ but only select $T_{q_2}^{(2)} + T_{q_3}^{(3)}$ terms contribute. \cite{Andrews1978, Andrews1990}
%there is some fancy way to note that theres no index constraint but i dont remember the language...
As such, its decomposition into irreducible tensor form is allowed only if the values of $q', q''$ are constrained.
By substituting the spherical tensor definitions of $\vec{\mu}$ and $\beta_{ijk}$ into \autoref{sivebeta}, it is clear that only $\Delta J = \pm 1$ transitions are allowed, i.e., SIVE-IR features arise only from P or R branches (assuming the ground state is $\ket{g00}$).
This is expected since $\vec{\mu}$ transitions demand $\Delta J = \pm 1$.
As a result, consistent with vibrational SR-SIVE, rovibrational SR-SIVE follows the same rules as common infrared rovibrational spectroscopy.

\subsection{Quantitative SR-SIVE}
With the selection rules of SR-SIVE understood for vibrational and rovibrational spectroscopy, it becomes possible to extract quantitative information from SIVE spectra to compare to incoherent methods and other CMDS methods.

Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re{\chi^{(3)}}$ and $\Im{\chi^{(3)}}$.\cite{Levenson1974_1, Levenson1974_2}
Resonant SIVE lineshapes are complicated by amplitude level interference between the sample and its substrate and/or sample cell windows. \cite{RN362}
These effects have been treated in excellent detail for three laser infrared-infrared-visible SIVE using the pathway in \autoref{fig:sivewmel}a, and provide methods to export quantitative information from SIVE spectra. \cite{RN418}

Most quantitative methods in an n$^{th}$ order CMDS experiment are used to extract $\chi^{(n)}$ values for comparing the relative strength of nonlinear processes in different media. \cite{Zhu87, RN351, RN345}
The recorded $\chi^{(n)}$ values provide insight into how microscopic quantities ($\vec{\mu}, \alpha_{ij}, \beta_{ijk}$) impact nonlinear output.
To our knowledge, only 2D-IR spectroscopy has taken advantage of orientationally averaged $\chi^{(3)}$ values to extract transition moments.\cite{Grechko2012}
Previous treatments did not consider using quantitative analysis in SIVE to measure $\beta_{ijk}$ values from SIVE spectra.
It is thus useful to investigate how a simple treatment of orientational averaging can extract $\beta_{ijk}$ from a $\chi^{(3)}_{IJKL}$ expression.

The SIVE $\chi^{(3)}_{IJKL}$ expresson is
\begin{equation}\label{chi3}
\begin{split}
		\chi^{(3)}_{IJKL} &\sim \langle \gamma_{ijkl} \rangle \sim \langle \beta_{ijk} \mu_l \rangle\\
\end{split}
\end{equation}
The steps behind orientational averaging of $\gamma_{ijkl}$ are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Orientational averaging for specific input polarizations shows that specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. 
Infrared spectra can provide values for different $\vec{\mu}$ components.
Since $\vec{\mu}$ is a rank-one tensor, its three Cartesian components share the simple relationship $\langle {\mu^2_x} \rangle = \langle {\mu^2_y} \rangle = \langle {\mu^2_z} \rangle = \frac{1}{3}\langle {\mu^2} \rangle$. \cite{RN459}
Neglecting rotational dependence, $\langle {\mu^2} \rangle = \abs{\mu}^2 = \mu^2$.
Substituting into \autoref{chi3} shows that
\begin{subequations}
	\begin{equation}
		\begin{split}
			\chi^{(3)}_{IJKL} &\sim \langle \beta_{ijk} \mu_l \rangle\\
			&\sim \langle \beta_{ijk} \rangle \frac{\mu}{\sqrt{3}} 
		\end{split}
	\end{equation}
This is done because any element of $\vec{\mu}$ is easily related to its magnitude through the directional cosine relationship shown above. 
Continuing, we see
	\begin{equation}
		\langle \beta_{ijk} \rangle \sim \frac{\sqrt{3}}{\mu} \chi^{(3)}_{IJKL}
	\end{equation}
\end{subequations}
Since $\mu^2$ can be extracted by fitting infrared spectra,\cite{RN412} it is clear that different polarization schemes in SIVE spectroscopy give a measurement of $\beta_{ijk}$ values.
SIVE can therefore provide useful insight into the sensitivity of hyper-Raman spectroscopy to different infrared active vibrations.

\subsection{Coherent Interference in SR-SIVE}
It is well known that nonlinear wave mixing processes which share the same time ordering and output frequency will interfere. \cite{RN135, Bonn2024}
Understanding interfering pathways is useful because they can, in some cases, eliminate output. \cite{RN342, RN135}
The WMEL diagrams presented in \autoref{fig:sivewmel} outline resonant SIVE processes when the infrared pulse $(\omega_1)$ is resonant with a vibrational mode. 
However, if $\omega_2$ becomes vibrationally resonant and $\omega_1$ is non-resonant, then other SIVE pathways appear.\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:sivewmel}, the WMEL diagrams presented in \autoref{fig:sivewmel2} may significantly interfere with DOVE-IR and DOVE-Raman spectra.
The time ordering in DOVE-IR and DOVE-Raman spectra is used to interpret anharmonicities, as DOVE-IR and DOVE-Raman destructively interfere when outputs are at the same frequency. \cite{RN135, RN324}
It is therefore important to understand how the corresponding SIVE pathways interfere.
To this end, we investigate how the SIVE-IR-II and SIVE-Raman pathways (\autoref{fig:sivewmel2}a, b, respectively) interfere under the rotating wave approximation.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{main/wmel2.png}
	\caption{Wave Mixing Energy Level (WMEL) diagrams of interfering SR-SIVE pathways in the  $\vec{k}_4 = -\vec{k}_1 + \vec{k}_2 + \vec{k}_3$ phasematching geometry. 
	(a) is similar to the process diagrammed in \autoref{fig:sivewmel}, whereas (b) is similar to a singly resonant coherent anti-Stokes Raman-type process.}
	\label{fig:sivewmel2}
\end{figure}
\begin{widetext}
We employ the response function formalism of Mukamel for this analysis. \cite{RN287}
The third order response functions, $R^{(3)}$, for SR-SIVE-IR-II ($R^{(3)}_{S_{II}}$) and SR-SIVE-Raman ($R^{(3)}_{R}$) (\autoref{fig:sivewmel2}a, b) are 
\begin{subequations}
	\begin{equation} \label{mixing:a}
		R^{(3)}_{S_{II}} (t_2) = -\frac{i}{\hbar} [\beta(t_2), \mu(0)] \exp(-i\Delta_{v''g}t_2) \theta(t_2)
	\end{equation}
	\begin{equation}\label{mixing:b}
		\begin{split}
			R^{(3)}_{R} (t_2) & = \frac{i}{\hbar} [\beta(t_2), \mu(0)] \exp(-i\Delta_{v''g}t_2) \theta(t_2) = - R^{(3)}_{S_{II}} (t_2)\\
		\end{split}
	\end{equation}
\end{subequations}
By assuming that the beams which stimulate hyper-Raman scattering, $\vec{E}_1, \vec{E}_3$, interact simultaneously at some time $\tau$ after $\vec{E}_2$ interacts with the sample, the total output polarization, $\vec{P}^{(3)}(t)$, is written as \cite{Cho2001}
	\begin{equation}\label{polarization_SRSIVE_a}
		\begin{split}
			\vec{P}^{(3)}(t) &= \vec{P}_{S_{II}}(t) + \vec{P}_{R}(t)\\
			&= \int_0^\infty \mathrm{d}t_2 (R^{(3)}_{S_{II}} (t_2) + R^{(3)}_{R} (t_2))\vdots \vec{E}_3(t) \vec{E}_1(t) \vec{E}_2(t-t_2 + \tau)\\
			&= \int_0^\infty \mathrm{d}t_2 (R^{(3)}_{S_{II}} (t_2) - R^{(3)}_{S_{II}} (t_2))\vdots \vec{E}_3(t) \vec{E}_1(t) \vec{E}_2(t-t_2 + \tau) \\
			&= 0
		\end{split}
	\end{equation}
It is clear the total output polarization for SR-SIVE-IR-II and SR-SIVE-Raman vanishes, or, SR-SIVE-IR-II and SR-SIVE-Raman perfectly destructively interfere.
\end{widetext}
The absence of SR-SIVE-IR-II pathways in ultrafast DOVE work is rationalized by this calculation. \cite{RN367, McDonnell2024}
In the case of the time ordered SR-SIVE-IR-I pathway, there are no opportunities for interference as all other pathways are nonresonant with the first interaction. 
A possible mechanism for interfering output in the SR-SIVE-I pathway are cascaded second order processes (SFG, DFG). \cite{RN243, RN300}
In isotropic media, SFG and DFG vanish under the electric dipole approximation,\cite{Zhu87} so cascades are unlikely to inhibit output in centrosymmetric media.
In media where SFG and DFG are allowed, e.g. surfaces or noncentrosymmetric media, second order cascades may become important. 
Generally, for nonlinear spectroscopies which share the same phasematching conditions as SR-SIVE (e.g., DOVE), it is useful to note that SR-SIVE-IR-I is not canceled, whereas SR-SIVE-IR-II and SR-SIVE-Raman permanently cancel.
